---
title: "OIDD245: Data Project 2 - Analytics for Good"
output: html_notebook
---

```{r}
install.packages("sf")
install.packages("corrplot")
install.packages("ggcorrplot")
install.packages("ggmap")
install.packages("choroplethr")
install.packages("choroplethrMaps")

# import necessay libraries
library(tidyverse)
library(lubridate)
library(ggplot2)
library(dplyr)
library(tidyr)
library(sf)
library(ggcorrplot)
library(choroplethr)
library(ggmap)
```
# Question I want to answer:
1. What are the happiest countries in the world?
2. What factors affects this happiness score?
3. What countries have the highest number of internet users?
4. How is access to internet and happiness related?

Main Q: What changes can countries make to make their people happier?
```{r}
# read in csv files 
internet_data = read.csv("gapminder_internet.csv", stringsAsFactors=FALSE)
happiness_data = read.csv("world-happiness-report.csv", stringsAsFactors=FALSE)
happiness2021_data = read.csv("world-happiness-report-2021.csv", stringsAsFactors=FALSE)
```

```{r}
# number of countries' data we have in each year 2005-2020
count(happiness_data, happiness_data$year, sort=TRUE)
happiness2021_data
```

```{r}
happiness2021_data$Country.name = toString(happiness2021_data$Country.name)
```

```{r}
# 2021 top 10 happiest countries
top_10_2021 <- head(happiness2021_data[order(-happiness2021_data$Ladder.score),], n=10)
top_10_2021

ggplot(data=top_10_2021, aes(x=reorder(Country.name, Ladder.score), y=Ladder.score)) +
  geom_bar(stat="identity", fill="#FDEE87") + coord_flip() + theme_linedraw() + ggtitle("2021 Top 10 Happiest Countries") + xlab("Countries") + ylab("Ladder Score")
```
```{r}
# 2021 bottom 10 happiest countries
last_10_2021 <- tail(happiness2021_data[order(-happiness2021_data$Ladder.score),], n=10)
last_10_2021
```

```{r}
# 2020 top 10 countries
data_2020 =  happiness_data[which(happiness_data$year==2020),]
top_10_2020 <- head(data_2020[order(-data_2020$Life.Ladder),], n=10)
top_10_2020

ggplot(data=top_10_2020, aes(x=reorder(Country.name, Life.Ladder), y=Life.Ladder)) + geom_bar(stat="identity", fill="#FDEE87") + coord_flip() + theme_linedraw() + ggtitle("2020 Top 10 Happiest Countries") + xlab("Countries") + ylab("Ladder Score")
```

```{r}
# 2019 top 10 countries
data_2019 =  happiness_data[which(happiness_data$year==2019),]
top_10_2019 <- head(data_2019[order(-data_2019$Life.Ladder),], n=10)
top_10_2019

ggplot(data=top_10_2019, aes(x=reorder(Country.name, Life.Ladder), y=Life.Ladder)) + geom_bar(stat="identity", fill="#FDEE87") + coord_flip() + theme_linedraw() + ggtitle("2019 Top 10 Happiest Countries") + xlab("Countries") + ylab("Ladder Score")
```

```{r}
# 2018 top 10 countries
data_2018 =  happiness_data[which(happiness_data$year==2018),]
top_10_2018 <- head(data_2018[order(-data_2018$Life.Ladder),], n=10)
top_10_2018

ggplot(data=top_10_2018, aes(x=reorder(Country.name, Life.Ladder), y=Life.Ladder)) + geom_bar(stat="identity", fill="#FDEE87") + coord_flip() + theme_linedraw() + ggtitle("2018 Top 10 Happiest Countries") + xlab("Countries") + ylab("Ladder Score")
```

```{r}
# 2005 top 10 countries
data_2005 = happiness_data[which(happiness_data$year==2005),]
top_10_2005 <- head(data_2005[order(-data_2005$Life.Ladder),], n=10)
top_10_2005

ggplot(data=top_10_2005, aes(x=reorder(Country.name, Life.Ladder), y=Life.Ladder)) + geom_bar(stat="identity", fill="#FDEE87") + coord_flip() + theme_linedraw() + ggtitle("2005 Top 10 Happiest Countries") + xlab("Countries") + ylab("Ladder Score")
```

As we can see Finland not only tops in 2021, but has been the leading country in terms of happiness for the past three years.

Why makes Finland have the highest happiness score? 

Let's start by looking at the logged GDP? Does money actually equal happiness?
```{r}
finland = happiness2021_data[which(happiness2021_data$Country.name=="Finland"),]
finland$Logged.GDP.per.capita
afghanistan = happiness2021_data[which(happiness2021_data$Country.name=="Afghanistan"),]
afghanistan$Logged.GDP.per.capita
#count(happiness2021_data, happiness2021_data$Logged.GDP.per.capita, sort=TRUE)

ggplot(happiness2021_data, aes(x=Logged.GDP.per.capita)) + 
  geom_histogram(binwidth=.2, fill="#50C878") + ggtitle("2021 Logged GDP") + xlab("Logged GDP per capita") + ylab("Country Count") + geom_vline(data=finland, aes(xintercept=Logged.GDP.per.capita, color="#E13F2A"),linetype="solid") + geom_vline(data=afghanistan, aes(xintercept=Logged.GDP.per.capita, color="#3FE0D0"),linetype="solid") +scale_color_discrete(name = "Legend", labels = c("Afghanistan", "Finland")) + theme_linedraw() 
```

```{r}
ggplot(happiness2021_data, aes(x=Social.support)) + 
  geom_histogram(binwidth=.025, fill="#50C878") + ggtitle("2021 Social Support") + xlab("Social Support") + ylab("Country Count") + geom_vline(data=finland, aes(xintercept=Social.support, color="#E13F2A"),linetype="solid") + geom_vline(data=afghanistan, aes(xintercept=Social.support, color="#3FE0D0"),linetype="solid") +scale_color_discrete(name = "Legend", labels = c("Afghanistan", "Finland")) +theme_linedraw() 
```
```{r}
ggplot(happiness2021_data, aes(x=Healthy.life.expectancy)) + 
  geom_histogram(binwidth=1, fill="#50C878") + ggtitle("2021 Healthy Life Expectancy") + xlab("Healthy Life Expectancy") + ylab("Country Count") + geom_vline(data=finland, aes(xintercept=Healthy.life.expectancy, color="#E13F2A"),linetype="solid") + geom_vline(data=afghanistan, aes(xintercept=Healthy.life.expectancy, color="#3FE0D0"),linetype="solid") +scale_color_discrete(name = "Legend", labels = c("Afghanistan", "Finland")) +theme_linedraw() 
```

```{r}
ggplot(happiness2021_data, aes(x=Freedom.to.make.life.choices)) + 
  geom_histogram(binwidth=.025, fill="#50C878") + ggtitle("2021 Freedom to Make Life Choices") + xlab("Freedom to Make Life Choices") + ylab("Country Count") + geom_vline(data=finland, aes(xintercept=Freedom.to.make.life.choices, color="#E13F2A"),linetype="solid") + geom_vline(data=afghanistan, aes(xintercept=Freedom.to.make.life.choices, color="#3FE0D0"),linetype="solid") +scale_color_discrete(name = "Legend", labels = c("Afghanistan", "Finland")) +theme_linedraw() 
```

```{r}
ggplot(happiness2021_data, aes(x=Generosity)) + 
  geom_histogram(binwidth=.025, fill="#50C878") + ggtitle("2021 Generosity") + xlab("Generosity") + ylab("Country Count") + geom_vline(data=finland, aes(xintercept=Generosity, color="#E13F2A"),linetype="solid") + geom_vline(data=afghanistan, aes(xintercept=Generosity, color="#3FE0D0"),linetype="solid") +scale_color_discrete(name = "Legend", labels = c("Afghanistan", "Finland")) +theme_linedraw() 
```

```{r}
ggplot(happiness2021_data, aes(x=Perceptions.of.corruption)) + 
  geom_histogram(binwidth=.025, fill="#50C878") + ggtitle("2021 Perceptions of Corruption") + xlab("Perceptions of Corruption") + ylab("Country Count") + geom_vline(data=finland, aes(xintercept=Perceptions.of.corruption, color="#E13F2A"),linetype="solid") + geom_vline(data=afghanistan, aes(xintercept=Perceptions.of.corruption, color="#3FE0D0"),linetype="solid") +scale_color_discrete(name = "Legend", labels = c("Afghanistan", "Finland")) +theme_linedraw() 
```

```{r}
#grouby_region = df.groupby('Regional indicator')['Ladder score'].mean().reset_index().sort_values('Ladder score')
grouby_region <- happiness2021_data %>% group_by(Regional.indicator) %>%
	summarise(mean_score = mean(Ladder.score))
grouby_region <- grouby_region[order(-grouby_region$mean_score),]

grouby_region

ggplot(data=grouby_region, aes(x=reorder(Regional.indicator, mean_score), y=mean_score)) + geom_bar(stat="identity", fill="#FDEE87") + coord_flip() + theme_linedraw() + ggtitle("Regions Ranked by Happiness") + xlab("Regions") + ylab("Avg. Ladder Score")
```
```{r}
corr_data <- happiness2021_data %>% select(Ladder.score, Logged.GDP.per.capita, Social.support, Healthy.life.expectancy, Freedom.to.make.life.choices, Generosity, Perceptions.of.corruption)
corr_data

corr <- round(cor(corr_data), 1)
corr
ggcorrplot(corr, ggtheme = ggplot2::theme_minimal(),
   colors = c("#FDEE87", "white", "#50C878"), lab=TRUE) 
ggcorrplot(corr, ggtheme = ggplot2::theme_minimal(),
   colors = c("#FDEE87", "white", "#50C878")) 
```
```{r}
happiness2021_data$Country.name
```

```{r}
plotdata <- happiness2021_data %>%
  rename(region = Country.name,
         value = Ladder.score) %>%
  mutate(region = tolower(region)) %>%
  mutate(region = recode(region,
                          "taiwan province of china" = "taiwan",
                          "hong kong s.a.r. of china" = "hong kong",
                          "congo (brazzaville)" = "republic of congo",
                          "palestinian territories" = "palestine",
                          "united states" = "united states of america",
                          "tanzania" = "united republic of tanzania",
                          "north cyprus" = "northern cyprus",
                          "serbia" = "republic of serbia"))
country_choropleth(plotdata) + 
  scale_fill_brewer(palette="YlGn") +
  labs(title = "2021 Ladder Score by Country",
       caption = "source: world happiness report 2021",
       fill = "Ladder Score")
```

```{r}
plotinternetdata <- internet_data %>%
  rename(region = country,
         value = internetuserate) %>%
  mutate(region = tolower(region)) %>%
  mutate(region = recode(region,
                          "united states" = "united states of america",
                          "congo, rep." = "republic of congo",
                          "congo, dem. rep." = "democratic republic of the congo",
                          "korea, dem. rep." = "south korea",
                          "korea, rep." = "north korea",
                          "dominican rep." = "dominican repulic",
                          "czech rep." = "czech republic",
                          "cote d'ivoire" = "ivory coast",
                          "bahamas" = "the bahamas",
                          "dominican repulic" = "dominican republic",
                          "tanzania" = "united republic of tanzania",
                          "guinea-bissau" = "guinea bissau"))
country_choropleth(plotinternetdata) + 
  scale_fill_brewer(palette="YlGnBu") +
  labs(title = "Internet User Rate by Country",
       caption = "source: Gapminder",
       fill = "Internet User Rate")
```

```{r}
# remove NA entries

corr_internetdata <- na.omit(internet_data) %>% select(incomeperperson, internetuserate, urbanrate)
corr_internetdata

corr_internet <- round(cor(corr_internetdata), 1)
corr_internet
ggcorrplot(corr_internet, ggtheme = ggplot2::theme_minimal,
   colors = c("#FDEE87", "white", "#3FE0D0"), lab=TRUE)
ggcorrplot(corr_internet, ggtheme = ggplot2::theme_minimal,
   colors = c("#FDEE87", "white", "#3FE0D0"))
```

```{r}
# join datasets to see correlation with ladder score
internet_data = internet_data %>% 
  rename(Country.name = country)
joined_data <- merge(happiness2021_data, internet_data, by="Country.name")
joined_data

corr_internetdata <- na.omit(joined_data) %>% select(Ladder.score, incomeperperson, internetuserate, urbanrate)
corr_internetdata

corr_internet <- round(cor(corr_internetdata), 1)
corr_internet
ggcorrplot(corr_internet, ggtheme = ggplot2::theme_minimal,
   colors = c("#FDEE87", "white", "#3FE0D0"), lab=TRUE)
ggcorrplot(corr_internet, ggtheme = ggplot2::theme_minimal,
   colors = c("#FDEE87", "white", "#3FE0D0"))
```

```{r}
# top 10 internet user countries
top_10_internet <- head(internet_data[order(-internet_data$internetuserate),], n=10)
top_10_internet

ggplot(data=top_10_internet, aes(x=reorder(Country.name, internetuserate), y=internetuserate)) + geom_bar(stat="identity", fill="#3FE0D0") + theme_linedraw() + ggtitle("Top 10 Internet Usage Countries") + xlab("Countries") + ylab("Internet Use Rate") + theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5)) 
```

```{r}
# top 10 happiest countries ranked by internet user rate
top_10_happiest_internet <- head(joined_data[order(-joined_data$Ladder.score),], n=10)
top_10_happiest_internet <- top_10_happiest_internet[order(-top_10_happiest_internet$internetuserate),] 

ggplot(data=top_10_happiest_internet, aes(x=reorder(Country.name, internetuserate), y=internetuserate)) + geom_bar(stat="identity", fill="#3FE0D0") + theme_classic() + ggtitle("2021 Top 10 Happiest Countries by Internet User Rate") + xlab("Countries") + ylab("Internet Use Rate") + theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5)) 
```
```{r}
# ladder score vs internet user rate
ggplot(joined_data, aes(x=Ladder.score, y=internetuserate)) + 
  geom_point(color="#3FE0D0")+
  geom_smooth(method=lm, se=FALSE, color="#FDEE87") + theme_linedraw() + ggtitle("Ladder Score vs Internet Use Rate") + xlab("Ladder Score") + ylab("Internet Use Rate")
```
```{r}
# internet user rate vs urban rate
ggplot(joined_data, aes(x=urbanrate, y=internetuserate)) + 
  geom_point(color="#3FE0D0")+
  geom_smooth(method=lm, se=FALSE, color="#FDEE87") + theme_linedraw() + ggtitle("Urban Rate vs Internet Use Rate") + xlab("Urban Rate") + ylab("Internet User Rate")
```

```{r}
# income per person vs internet user rate
ggplot(joined_data, aes(x=incomeperperson, y=internetuserate)) + 
  geom_point(color="#3FE0D0")+
  geom_smooth(method=lm, se=FALSE, color="#FDEE87") + theme_linedraw() + ggtitle("Income Per Person vs Internet Use Rate") + xlab("Income per Person") + ylab("Internet Use Rate")
```

By: Joey Zhao
